### 2005-12-13 inst/paper/paper2.tex
Updated paper 2.
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MVCClass@15362 bc3139a8-67e5-0310-9ffc-ced21a209358

\documentclass{article}[11pt]
\setlength{\topmargin}{-2cm}
\setlength{\textheight}{24cm}
\setlength{\oddsidemargin}{10mm}
\setlength{\evensidemargin}{10mm}
\setlength{\textwidth}{15cm}
\setlength{\parskip}{3mm}
\setlength{\parindent}{0mm}

\usepackage{epsfig}
\usepackage{comment}
\usepackage{graphics}
\usepackage{graphicx}

\newcommand{\Rfunction}[1]{{\texttt{#1}}}
\newcommand{\Robject}[1]{{\texttt{#1}}}
\newcommand{\Rpackage}[1]{{\textit{#1}}}
\newcommand{\Rslot}[1]{\textsl{#1}}

\begin{document}

\bibliographystyle{amsplain}
\bibliography{paper2}

\begin{titlepage}
\begin{center}
\Large
\vspace*{1in}
\textbf{Creating Linked, Interactive Views to \linebreak Explore Linked Data} \\
\vspace{1in}
\textbf{Dana-Farber Cancer Institute}\\ 
\textbf{December 9, 2005} \\
\vspace{1in} 
\textbf{Elizabeth Whalen} \\
\end{center}
\end{titlepage}
\normalsize
\newpage

%\renewcommand{\baselinestretch}{1.5}
\section{Introduction}\label{Sec:Intro}

\subsection{Background}\label{Ssec:Backg}

There are many ways to visualize multivariate data where more than
two or three dimensions are needed to view the data.  Several of these ideas
were discussed in a previous paper, titled ``Creating Linked, Interactive
Views to Explore Multivariate Data'' \cite{EW05}, so the focus in this paper
will be one specific method of visualizing multivariate data that is
particularly useful for exploratory data analysis.  This paper will discuss
creating linked, interactive views of linked data sets.

First brief definitions of linked views, interactive views and linked data
sets will be given.  Linked views mean that if a component that represents
data (such as a point on a plot or a row in a spreadsheet) changes its
appearance on one view, then the corresponding component on a second view will
also change its appearance.  For components on different views to be
corresponding, the data displayed in those components must come from the same
entity in a model.  As an example, suppose that the model consisted of four
variables for the fifty states' murder rate, assault rate, rape rate, and
percent of population living in urban areas.  Now the model could be
represented as a matrix of four columns by fifty rows with four columns for
the four variables and fifty rows for the fifty states.  In this model, the
entities would be the states.  If two scatterplots were created as views of
this model, where the first plot displayed murder rate versus assault rate and
the second plot displayed rape rate versus percent urban population, then the
points on each plot that referred to the state of Montana would be
corresponding components.

Interactivity is an idea that will be familiar from using the Internet.  For
example, clicking on a hyperlink on a web page results in a new web page
opening.  Another example is clicking on the login button after entering a
user name and password to view one's email.  Interactivity means that there is
a response to some user action.

One particular model for creating interactive applications involves three
components: some user action or input, which will be referred to as an event;
a response to that action, which will be executed by a function that is called
a callback function; and a method that connects the event to the callback
function, which will be referred to as the signal handler.  The events can
range from the user moving the cursor over a plot to pressing a button on a
graphical user interface (GUI) to pressing certain keys.  Notice that the
events almost always occur over some type of interface, though some events,
such as key press events may occur without an interface.  The callback
function is a unique function because its signature is determined by the event
that the callback function is linked to.  The parameters of the callback
function may include information about the event, such as the location where
the event occurred, and information about the object that just caused the
event, such as the button on the GUI.  The signal handler is the piece that
must look to see if an event has occurred and then when the event has
occurred, the signal handler calls the appropriate callback function.  The
strengths of this model of interactivity are that programmers can decide which
events to notice by deciding which signal handlers to define, programmers
can change the response to an event by changing the callback function that is
linked to an event, and the system's event loop can continue working on other
things until a signal is emitted because an event occurred.

These three pieces in this model of interactivity are ordered as follows:
an event occurs, the event causes a signal to be emitted that is caught by the
signal handler, and then the signal handler calls the callback function, which
results in the response to the user action.  A flowchart of these steps is
shown below in Figure \ref{Fig:Event}.  In the example of clicking on a
hyperlink mentioned above, the event is the user clicking on the
hyperlink and the callback function causes a new web page to open.
The signal handler, which is the method that connects the event to the
callback function, is not noticed by the user, but it is handled by the
system's event loop when a signal is emitted.

\clearpage

\begin{figure}[!h]
  \begin{center}    
    \scalebox{0.6}{\includegraphics{Event.jpg}}
    \caption{ Flow Chart of the Response to an Event. }
    \label{Fig:Event}
  \end{center}
\end{figure}

Linked data sets mean that the data have more than one conceptual grouping,
but there is a relationship between these different conceptual groupings.
Linked data sets are an idea that will be familiar to users of relational
databases.  Data in a relational database are typically stored in many
linked tables with rows in different tables linked to each other through a
unique identifier called a key.  In this example, each table would be
considered a data set and they are linked through the keys.  Another situation
where linked data sets occur is when there is experimental or study data that
are linked to meta-data.  An example of this situation is given in Section
\ref{Ssec:Limit} where the experimental data are microarray data, which looks
at gene expression levels, and the meta-data are a Gene Ontology graph that
gives the molecular functions of the genes studied in the microarray
experiment.

Creating linked, interactive views of data is flexible for exploring
multivariate data because many different views can be created, such as
scatterplots, spreadsheets, and heatmaps, and these views are linked because
they are based on the same underlying model.  Thus, users can decide which
views best visually represent their data.  Also, since the views are
interactive, users can change the views while looking at them to make the
views more informative about the underlying data structure.  This flexibility
in linked, interactive views gives users a powerful tool for visually
exploring their data.

In Section \ref{Ssec:Design}, the design of linked, interactive views based on
the model-view-controller (MVC) paradigm will be discussed.  However, the
limitation with this design is that it expects one data set, which limits the
type of data we can visualize.  There are many
situations where multiple data sets occur, particularly any time there are
experimental or study data linked to meta-data.  Being able to visualize data
from multiple data sets would be helpful to understand the relationships
between variables.  Thus, after discussing the limitations of the MVC design
in Section \ref{Ssec:Limit}, an extension of the MVC design will be discussed
in Section \ref{Sec:Extend}.  This extension will allow the creation of linked,
interactive views based on linked data sets.  The software that is discussed
in this paper will have three goals: to create linked views of linked data
sets, to create interactive views where the response to an event can be
changed, and to create an extensible design that will allow future users to
make additions.

As a motivating example of linked views based on linked data sets, consider
Figure \ref{Fig:firstMP}.  In this figure, there are two views, one view of
data set 1 and one view of data set 2.  Data set 1 and data set 2 are linked.
When a user initially interacts with the scatterplot view of data set 2, the
response to this interaction is to tell data set 2 to update itself (this is
indicated on the figure as a circled 1).  Then in step 2, data set 2 tells the
scatterplot view that is should update itself in response to the change in the
data.  Next, in step 3, because data set 2 has changed and because it is
linked to data set 1, data set 2 sends information to data set 1 about the
change.  In step 4, the information from data set 2 is converted into
information that data set 1 can understand.  In step 5, this information is
received by data set 1 and data set 1 updates its data.  Finally, in step 6,
data set 1 tells its view to update itself.  Thus, in this example, the views
are synchronized based on the underlying linked data sets.

\clearpage

\begin{figure}[ht]
  \begin{center}
    \scalebox{0.6}{\includegraphics{firstMessagePassing.jpg}}
    \caption{ Linked Views of Linked Data Sets. }
    \label{Fig:firstMP}
  \end{center}
\end{figure}

\subsection{Design and Implementation of Linked, Interactive
  Views}\label{Ssec:Design}

Focusing on linked, interactive views as a powerful visualization tool, the
next step is to determine how to design a software package that can implement
this method.  A popular design for creating linked views of data is the
model-view-controller (MVC) paradigm.  This design consists of three types
of objects: the controller, which defines what actions occur in response to
user input; the view, which consists of displays of the data; and the model,
which manages the data.  This paradigm is powerful because it decouples the
views from the model.  Thus, only one copy of the data needs to be stored and
then all views refer to this data set.  With all views based on the same data,
any changes to the data can be propagated to all views, which will create
linked views.

To link the views to this one data set there is a subscribe/notify procedure
between the views and the model.  The views subscribe to a particular model
and the model must notify the views when a change occurs so that the views
will be updated.  This design allows multiple views of the model that are
linked, but are not aware of each other, as shown in Figure \ref{Fig:ExMVC}
\cite{DesignPatterns}.  Each view depends only on visually representing the
model and thus, it is not necessary for a view to be aware of other views.  By
keeping each view ignorant of other views, other views can added and removed
without affecting a different view.

\begin{figure}[ht]
  \begin{center}
    \scalebox{0.6}{\includegraphics{ExMVC.jpg}}
    \caption{ A Model with its Views. }
    \label{Fig:ExMVC}
  \end{center}
\end{figure}

Notice that although the components (model, view, and controller) in the MVC
design are independent of one other, there still needs to be communication
between them so that information can be shared.  Examples of this
communication between components include the model notifying views that the
data have changed, a view asking the controller how it should respond to user
interaction, and a view telling the model that the data should change in
response to user interaction with the view.  This message passing between
components in the MVC paradigm will be discussed in detail in Section
\ref{Ssec:OneMess}.

The ``Creating Linked, Interactive Views to Explore Multivariate Data''
\cite{EW05} paper discussed the creation of R software to fulfill three goals:
to create views that are linked, to create views that are interactive, and to
create a design that is extensible.  To satisfy the overall goals of the
software, two R packages, called \Rpackage{MVCClass} and \Rpackage{iSPlot},
were created.  The \Rpackage{MVCClass} package implemented the class and
generic function definitions needed for the MVC design and the
\Rpackage{iSPlot} package created the front end to let the user create linked,
interactive views of two-dimensional data, such as data frames and matrices,
through either a GUI or through command line functions.  The reason for
separating the software into two packages was that it allowed the class
definitions in the \Rpackage{MVCClass} package to be used by any other R
package that would like to create linked views.

\subsection{Limitations of the MVC Design}\label{Ssec:Limit}

The MVC design assumes that there is one data set that all views will
reference.  But what if the data consisted of more than one conceptual
grouping?  For example, consider data stored in a relational database, where
there are multiple tables of data and rows in different tables are linked
through a unique identifier called a key.  Although these tables of data could
be combined into one large table, there are several reasons not to do this.
Keeping the data in separate tables reduces redundancy, particularly when the
rows are related by a one-to-many or a many-to-many relationship.  Also
separate data tables generally lead to more maintainable data.  By storing
only one copy of each piece of information (in a row in one table), it is
easier to ensure that the data are correctly updated when a change occurs.
All other tables that are connected with this information can then just
reference this row that has been updated.

Consider Figure \ref{Fig:DBTab} shown below.  Here the database consists of
three tables, State, City and StateCity.  The State table has three columns,
State (the key), Life Expectancy, and Income.  The City table has three
columns, City (the key), Precipitation, and Temperature.  The StateCity table
has two columns, State and City.  This last table relates which cities are in
which states.  In this example, each city can be in only one state, but each
state can have more than one city so there is a one-to-many relationship
between states and cities.

%\clearpage

\begin{figure}[ht]
  \begin{center}
    \scalebox{0.5}{\includegraphics{newDatabaseTables.jpg}}
    \caption{ Example of Relational Database Tables. }
    \label{Fig:DBTab}
  \end{center}
\end{figure}

Clearly, this information could be combined into one large table, called
CombinedStateCity, as shown in Figure \ref{Fig:OneDBTab}.  The
CombinedStateCity table would have the following columns: State, Life
Expectancy, Income, City, Precipitation, and Temperature.  Now if a state's
life expectancy changed, all instances of this state would need to be changed
in the new large table.  For example, if the state had the information for
three cities recorded, then the state would have three rows in the
CombinedStateCity table and each row would need to change the value for the
state's life expectancy.  If instead only one record was kept per state, as
shown in Figure \ref{Fig:DBTab} above, then this process would be less error
prone and would require less memory to store the data.  This example shows
that there are advantages to storing the data in separate structures.

%\clearpage

\begin{figure}[ht]
  \begin{center}
    \scalebox{0.5}{\includegraphics{newOneDatabaseTable.jpg}}
    \caption{ Example of a Combined Table. }
    \label{Fig:OneDBTab}
  \end{center}
\end{figure}

In a relational database all of the data are stored in tables, but consider the
situation where there are multiple data sets that are of different data
structures.  For example, suppose that you have experimental microarray data,
which look at gene expression levels.  This experimental data can be
stored as a two dimensional data structure, such as a matrix, where rows are
genes and columns are samples.  Linked to this experimental data are meta-data
that give information about the genes studied in the microarray experiment.
For example, this meta-data may be a Gene Ontology graph that gives the
molecular functions of these genes.  These two data sets (the experimental
data and the meta-data) are linked through the genes.  Each gene can have more
than one molecular function and each molecular function can have more than one
gene annotated at that function so there is a many-to-many relationship
between the genes and the molecular functions.  Though these two data sets are
linked, the microarray data and the Gene Ontology data are different data
structures.  The experimental microarray data can be represented as a matrix
and the Gene Ontology meta-data can be represented as a graph.

Now for the original MVC design, where one data set is expected, it is not
clear how to combine these two data sets into one data set.  The meta-data
graph could be stored as a matrix, but a graph is a data structure that can be
well represented as an object with slots for nodes and edges, especially if
the graph is sparse.  As with the relational database tables, it seems
best in terms of data storage, alteration, and retrieval to keep the two data
sets separate.  Also, keeping the two data sets separate does not force a data
structure on the data that may not fit well.

\section{Extending the MVC Design}\label{Sec:Extend}

As discussed in Section \ref{Sec:Intro}, the MVC design expects a single
conceptual unit, such as a data frame or a matrix,
with all views referencing this one data set.  However, in Section
\ref{Ssec:Limit} several examples were shown where multiple data sets are
linked.  These examples included linked tables in relational databases and
experimental data linked to meta-data.  Being able to visualize variables
simultaneously from different data sets that are linked would be useful for
exploring relationships between these variables.  Thus, the next step will be
to determine how the MVC design can be revised to account for multiple data
sets.

\subsection{One Large Data Set}\label{Ssec:OneDS}

One possibility for dealing with multiple data sets within the MVC design is to
combine the multiple data sets into one large data set.  With this new
combined data, the original MVC design is appropriate for creating linked
views.  Thus, the strength of combining the data sets into one large data set
is that the MVC paradigm is already appropriately set up to deal with the
one large data set.

Unfortunately, as discussed in Section \ref{Ssec:Limit}, combining
multiple data sets into one large data set is not always appropriate or even
possible.  Even if the multiple data sets have the same data structure,
combining the data sets often results in redundant data and makes updating the
data more complicated.  In the situation where the data sets are not of the
same data structure, it may not be possible to combine the data sets into one
data structure that retains all of the original information.  Also, combining
the data into one data structure may give the appearance that some variables
have a relationship that does not exist.  For example, suppose that in a
relational database, table 1's rows are linked to table 2's rows and table 3's
rows are also linked to table 2's rows.  Just because table 1 and table 3 are
both linked to table 2 does not necessarily mean that there is a meaningful
relationship between the data in table 1 and table 3.  However, by combining
the data in these three tables, it would give the appearance that there is a
relationship between the data in table 1 and table 3.  Thus, another
way of linking views based on multiple data sets needs to be designed.

\subsection{Multiple MVC Objects}\label{Ssec:MMVC}

Because of the difficulty of combining multiple data sets, the MVC paradigm
will need to be altered to fit the data.  It has been mentioned that the MVC
design is intended to deal with one data set.  A more straightforward approach
may be to link MVCs.  In linked MVCs, each MVC can be coherent and the
interactions between MVCs can be encapsulated in a set of messages.  One MVC
object to handle each data set.  Thus, within each MVC object, one could
create linked views of that particular data set, which is the same as the
original MVC design.  Then between MVC objects a messaging system would need
to be designed so that a change in one data set would be sent to the other
linked data sets.

\small
\begin{tabular}[t]{ | l | r | r | }
  \hline
  State & Life Expectancy & Income \\ \hline
  Pennsylvania & 70.43 & 4449 \\ \hline
  California & 71.71 & 5114 \\ \hline
  Minnesota & 72.96 & 4675 \\ \hline
\end{tabular}
\hspace{10pt}
\begin{tabular}[t]{ | l | r | r | }
  \hline
  City & Precipitation & Temperature \\ \hline
  Philadelphia & 39.9 & 39 \\ \hline
  Pittsburgh & 36.2 & 37 \\ \hline
  Los Angeles & 14.0 & 68 \\ \hline
  San Francisco & 20.7 & 56 \\ \hline
  Sacramento & 17.2 & 55 \\ \hline
  Duluth & 30.2 & 18 \\ \hline
  Minneapolis St. Paul & 25.9 & 22 \\ \hline
\end{tabular}

\begin{table}[h]
  \begin{center}
    \begin{tabular}{ | r | r | }
      \hline
      State & City \\ \hline
      Pennsylvania & Philadelphia \\ \hline
      Pennsylvania & Pittsburgh \\ \hline
      California & Los Angeles \\ \hline
      California & San Francisco \\ \hline
      California & Sacramento \\ \hline
      Minnesota & Duluth \\ \hline
      Minnesota & Minneapolis St. Paul \\ \hline
    \end{tabular}
    \caption{State and City Values.}\label{Tab:CityState}
  \end{center}
\end{table}

\normalsize
As an example, suppose that one has two data sets: one with information at the
state level and one with information at the city level.  These tables are
shown in Table \ref{Tab:CityState}.  The State data include Life Expectancy
(in years from 1970) and Income (per capita from 1974) data
while the City data include Precipitation (in inches from 1975) and Average
High Temperature in January (in Fahrenheit from 2005)
data.  These two data sets are linked by the CityState table that links cities
with the states they reside in.  Now for these two data sets there would be
two MVC objects.  One MVC object for the State data and its views and one MVC
object for the City data and its views.

\begin{figure}[ht]
  \begin{center}
    \scalebox{0.7}{\includegraphics{MultipleMVC.jpg}}
    \caption{ Example of Two MVC Objects. }
    \label{Fig:MultMVC}
  \end{center}
\end{figure}

Figure \ref{Fig:MultMVC} shows one view for the State data and one view
for the City data.  You can see how the State and City data are linked based
on the coloring of the points in the plots.  The points referring to the state
of Pennsylvania are red, the points referring to California are blue, and the
points referring to Minnesota are green.  Even though there are two MVC
objects here, with one MVC object for each data set, it is still clear that
the views are linked through the data sets and that is shown in this example
through the coloring.

This solution uses the power and simplicity of the MVC design and expands it
to the multiple data set problem.  Now we can build on the original design,
which has been well tested, instead of starting a new design from scratch.
The strength of this multiple linked MVC design is that it uses MVC
components, which can be individually created and then linked.  Thus, these
components are reusable.  Also this multiple MVC design is structured to fit
with the linking seen in the multiple data sets.  Clearly, the design should
be built around the problem and multiple MVC objects with messaging between
the MVC objects fits the problem of creating linked views of multiple linked
data sets.

Two parts of this design will need some special thought: creating an
appropriate messaging system between MVC objects and deciding how the MVC
objects will be related.  In designing an appropriate messaging system between
MVC objects, consideration will be given to what information should pass
between MVC objects and to how messages will propagate between MVC objects.
In deciding the relationship between the MVC objects, consideration will be
given to how the multiple data sets are related.  Clearly, the relationship
between the MVC objects should reflect the relationship between the data
sets.  The two examples of linked data sets that have been discussed
previously were linked tables in a relational database, and experimental data
that are linked to meta-data.  In this paper, the focus will be on experimental
data that are linked to meta-data, as those are the data we are most
interested in visualizing.  In this situation, the experimental data can be
thought of as determining what meta-data are studied.  In the example of
microarray experiment data linked to Gene Ontology meta-data, only Gene
Ontology data that refer to genes in the microarray experiment will be
included in the meta-data.  Thus, there is a hierarchical relationship between
these data sets.  Therefore, the design of the linking between MVC objects
will be based on a hierarchical relationship between the data sets, where one
MVC object creates another MVC object.  These decisions will be discussed in
the following sections as the design for multiple MVC objects is explained.

\section{Design Details of One MVC Object}\label{Sec:OneMVC}

\subsection{Overview}\label{SSec:OneOver}

Implementing an extensible design to create linked, interactive views of
linked data sets will be done in the R language in two packages, called
\Rpackage{MVCClass} and \Rpackage{iSNetwork}.  The \Rpackage{MVCClass} package
will implement the design of the MVC classes and generic functions.  The
\Rpackage{MVCClass} package will define the individual model, view, and
controller classes as well as the \Robject{MVC} class that binds the
individual model, view, and controller classes together into one unit.  The
package will also define the message classes that will allow the components to
communicate with each other.  Thus, the \Rpackage{MVCClass} package will
contain all of the classes necessary for creating linked views.   The goal of
separating these class and generic function definitions into one package is
that they will be available for other R packages that implement linked views
of data, such as the \Rpackage{iSPlot} and the \Rpackage{iSNetwork} packages.

With the class definitions created in the \Rpackage{MVCClass} package, the
\Rpackage{iSNetwork} package can decide which classes from \Rpackage{MVCClass}
to use based on the overall goals of the combined software.  The
\Rpackage{iSNetwork} package will also determine how objects of these classes
will behave because the methods for the classes are defined in the
\Rpackage{iSNetwork} package.  In addition to implementing the behavior of
objects, the \Rpackage{iSNetwork} package will also create a graphical user
interface (GUI) so that users can load data, create views, and interact with
the views through menus on the user interface.  As an alternative to the GUI,
users will also have access to the same functionality through command line
functions.

To create the GUI, \Rpackage{iSNetwork} will use the R packages,
\Rpackage{RGtk} and \Rpackage{gtkDevice}.  \Rpackage{RGtk} is an R package
that allows one to interact with Gtk functions from the R interface.  Gtk is
an open-source X Window toolkit \cite{Gtk} for creating user interfaces.  The
\Rpackage{gtkDevice} package creates a Gtk device by turning a Gtk drawing
area into a device that acts like an R device, but still responds to events so
that the view can be interactive.  For more information about these packages,
please see the paper, ``Creating Linked, Interactive Views to Explore
Multivariate Data'' \cite{EW05}.

The next subsections will explain the class definitions and the methods that
pertain to a single model-view-controller object.  Thus, the model classes,
the view classes, the controller, the \Robject{MVC} class (which binds the
model, view and controller classes together), and the message classes that
pass information between the three components in one MVC object will all be
discussed in the following subsections.  All of the class definitions are
implemented in the \Rpackage{MVCClass} package and all of the methods are
defined in the \Rpackage{iSNetwork} package.  Message class definitions that
pertain to communication between the MVC objects will be discussed in Section
\ref{Sec:MultMVC}.

\subsection{Model}\label{Ssec:OneModel}

In the MVC design, the model component is the place where data are stored and
updated.  Thus, the model is a storage location that can perform certain
operations.  Information can be sent to the model to add, update, and remove
data and the model can send out messages that changes have occurred to the
data that the model stores.  Thus, the object model for the model classes will
represent both the slots that are needed to store data and the methods that
are needed to perform the operations of the model.

Figure \ref{Fig:Model} shows the inheritance structure for the model classes.
Here the virtual class, \Robject{gModel}, has the slots \Rslot{modelData},
\Rslot{linkData}, \Rslot{virtualData}, and \Rslot{modelName}.  These four
slots are common information that all models will need to store.  The
\Rslot{modelData} slot is the data set for this model, the \Rslot{linkData}
slot is a list of two functions, \Rfunction{toParent} and
\Rfunction{fromParent}, which let this model send information to its parent
model and receive information from its parent model, respectively (see Section
\ref{Ssec:MultLink}), the \Rslot{virtualData} slot is data pertaining to the
views that needs to be stored with the model so that it can be shown in all
views of this model, and the \Rslot{modelName} slot is the name of the model.

%\clearpage

\begin{figure}[ht]
  \begin{center}
    \scalebox{0.7}{\includegraphics{ModelClass.jpg}}
    \caption{ Inheritance for Model Objects. }
    \label{Fig:Model}
  \end{center}
\end{figure}

As for specific model classes, the classes that are used in the
\Rpackage{iSNetwork} package are \Robject{dfModel}, \Robject{graphModel},
and \Robject{exprModel}.  The \Robject{dfModel} class represents a model where
the \Rslot{modelData} slot has class data frame (or matrix), the
\Robject{graphModel} class represents a model where the \Rslot{modelData} slot
has class graph, and the \Robject{exprModel} class represents a model where the
\Rslot{modelData} slot has class exprSet.

As mentioned previously, the model classes must also perform certain
operations and these operations are reflected in the methods that are defined
for the classes.  All three model classes have an \Rfunction{updateModel}
method that is defined in the \Rpackage{iSNetwork} package.  This method is
called whenever the data need to be updated in response to a user interacting
with a view.  Thus, the \Rfunction{updateModel} method is actually updating the
\Rslot{virtualData} slot because it is this slot in the model classes that
ensures that all views are appropriately visually linked.

Although only three model classes are currently implemented, future users are
welcome to make additions to the model classes for new data structures they
would like to represent.  For example, a user may want to add a new model
class called \Robject{tsModel} to represent time series data.  To make this
addition, a user could extend the \Robject{gModel} class in his or her own
package.

Note that the model classes must also notify views when the data have changed
so that the views can be updated.  The communication between a model and its
views will be discussed in Section \ref{Ssec:OneMess}.

\subsection{Views}\label{Ssec:OneViews}

The view classes represent the visual depictions of the model.  All views will
need to store some common information and respond to certain events through
methods.  This consideration led to an object model where the different view
classes inherit from a view virtual class, called \Robject{genView}.  The
object model for the view classes is shown in Figure \ref{Fig:View}.

\begin{figure}[ht]
  \begin{center}
    \scalebox{0.7}{\includegraphics{newViewClass.jpg}}
    \caption{ Inheritance for View Objects. }
    \label{Fig:View}
  \end{center}
\end{figure}

The \Robject{genView} class has three slots, \Rslot{dataName}, \Rslot{win},
and \Rslot{winNum}.  \Rslot{dataName} is the name of the model that the view
displays, \Rslot{win} is the Gtk window object that holds the view, and
\Rslot{winNum} is the number of the window so that the window can be
identified in the Window menu on the GUI (see Section \ref{Ssec:OneCont}).
All views will need to know these three pieces of information so
\Robject{genView} will bind the view classes together.

As for specific view classes, the \Robject{spreadView} class will represent a
spreadsheet view of the data.  This view will only make sense for models that
are a two dimensional data structure, such as a matrix or a data frame.  The
information needed for this view is stored in the slot, \Rslot{clist}, which
is the Gtk spreadsheet object.

When creating a plot of a data set, there is a general class, called
\Robject{plotView}, that has the slots, \Rslot{plotDevice}, \Rslot{plotPar},
and \Rslot{drArea}, which store the device number of the plot, the plotting
parameters, and the Gtk drawing area object, respectively.  This class is a
virtual class as it represents a general plot.  The instantiable plot classes
are the specific plot classes, \Robject{sPlotView}, \Robject{graphView}, and
\Robject{heatmapView}, which inherit from the \Robject{plotView} class.

The \Robject{sPlotView} class represents a scatterplot view and it has the
extra slots, \Rslot{dfRows}, \Rslot{xvar}, and \Rslot{yvar}, where
\Rslot{dfRows} are the row names from the model that are shown in the plot,
\Rslot{xvar} is the variable name from the model that is shown as the x
variable in the plot, and \Rslot{yvar} is the variable name from the model
that is shown as the y variable in the plot.  The \Robject{graphView} class
represents a graph plot view and it contains no extra slots besides those in
the \Robject{plotView} class.  The \Robject{heatmapView} represents a heatmap
view of the data and it contains the slot, \Rslot{ordering}, which is a list
that is returned from the \Rfunction{heatmap} function to give information
about the ordering in the row and column dendrograms.  Note that some of these
views are only applicable for certain model types. For example, the
\Robject{graphView} class will only make sense for a view of a graph object,
which would be stored in the \Robject{graphModel} class.

Since all views are expected to be interactive, all
view classes have methods that correspond to user interaction, such as
\Rfunction{clickEvent}.  Recall from Section \ref{Ssec:Backg} that when an
event occurs Gtk emits a signal that is caught by the signal handler.  The
signal handler then calls the callback function and it is in the callback
function that the corresponding method is called.  For example, when the
``clicked'' event occurs on a scatterplot view, the callback function calls the
scatterplot's \Rfunction{clickEvent} method.  Using the information that is
passed to the \Rfunction{clickEvent} method in the \Rslot{where} parameter,
the method converts the Gtk pixel coordinates that indicate where the event
occurred into user coordinates and then into an object from the plot, such as
a point on the plot.

For \Robject{sPlotView}, the \Rfunction{clickEvent} method is called when Gtk
emits a signal that a mouse button has been pressed and the
\Rfunction{motionEvent} method is called when Gtk emits a signal that
the cursor has been moved over the view.  Both the \Rfunction{clickEvent} and
\Rfunction{motionEvent} methods for a \Robject{sPlotView} object will
determine which point on the plot was clicked or under the cursor,
respectively, using the data in the \Rslot{where} parameter.  For
\Robject{spreadView}, the \Rfunction{clickEvent} method is called when Gtk
emits a signal that a row in the spreadsheet has been selected or unselected
and the \Rfunction{clickEvent} method will determine which row was changed.
For \Robject{graphView}, the \Rfunction{clickEvent} method is called when Gtk
emits a signal that a mouse button has been pressed and the
\Rfunction{motionEvent} is called when Gtk emits a signal that the cursor has
been moved over the view.  Here, though, the \Rfunction{clickEvent} and
\Rfunction{motionEvent} methods will determine which node was clicked or under
the cursor, respectively, using the data in the \Rslot{where} parameter.

The \Robject{heatmapView} class is different from the other views in that it
has neither a \Rfunction{clickEvent} or a \Rfunction{motionEvent} method.  The
current implementation of the \Rfunction{heatmap} function does not return
information about where plot objects are placed on the device.  As a result,
it is not possible to convert the user coordinates of where the event occurred
into an object (such as the row and column coordinates of the heatmap) on
the view.  Thus, the \Robject{heatmapView} class is the only view that is not
currently interactive.

As with the model classes, the view classes need an update method and thus, all
views have an \Rfunction{updateView} method that is defined in the
\Rpackage{iSNetwork} package.  This method must take the
information that a model has just been updated and appropriately update the
view to reflect that change.  Each method will depend on how the view
represents the data so for a \Robject{spreadView} object the
\Rfunction{updateView} method may select or unselect a row on the spreadsheeet
while for a \Robject{sPlotView} object the \Rfunction{updateView} method may
change the appearance of a point on the plot.

Note that when updating a view in R we do not have access to objects on a
view.  Thus, a point on a plot is not an object that allows us to update its
attributes and then have that point redrawn to reflect its new attributes.
Instead in R we must currently draw over a portion of the view to update the
view.  However, if the method of creating R plots changed, the
\Rfunction{updateView} methods could be modified.

The inheritance structure of the view classes is intended to be extensible.
For instance, if users wanted to add a new plot view, say a density plot, they
could create a new view class that inherits from \Robject{plotView}, called
\Robject{dPlotView}.  This new class, \Robject{dPlotView}, would have slots
named \Rslot{dfRows} and \Rslot{col} to store which rows and which column,
respectively, were used to calculate the density shown in the view.

\subsection{Controller}\label{Ssec:OneCont}

The controller must perform two tasks: the first task is to store information
and the second task is to allow the user to determine the course of actions.
For the first function, which is implemented in the \Rpackage{MVCClass}
package, the controller is implemented as an environment that is stored in the
\Rslot{controller} slot of a \Robject{MVC} instance, which will be discussed in
more detail in Section \ref{Ssec:OneMVC}.  Although other options exist for
storing data, such as a list or a database, an environment is nice because in
R an environment is not copied each time a value within the environment
changes.  The controller in this first role will need to store information
pertaining to state, such as information about the appearance of the views in
a particular model-view-controller entity.

For the second function, which is implemented in the \Rpackage{iSNetwork}
package, the controller gives the user access to the application program
interface (API), either through the GUI or through command line functions.
Through the API the user can load data, create views, set the response to user
interaction with a view, quit the program and perform other functions.  On the
GUI, there are menus to let users decide which action will be taken next.  If
the user would rather not use the GUI, there are command line functions that
provide access to the same functionality that is available through the GUI.
The GUI is shown in Figure \ref{Fig:ContWin} and will be referred to as the
control window.

\begin{figure}[ht]
  \begin{center}
    \scalebox{0.5}{\includegraphics{ControlWindow.jpg}}
    \caption{ The Control Window. }
    \label{Fig:ContWin}
  \end{center}
\end{figure}

As shown in Figure \ref{Fig:ContWin}, the menus on the control window are
File, Edit, Display, Event, and Window.  These menus always appear on the
control window, but the menu items under each menu will change depending on
the type of model that is active.  Because the goal is to have multiple
linked MVC objects, there will be more than one MVC object loaded and thus,
more than one model. However, the control window menus and the command line
functions will only perform operations on one MVC object at a time.  Thus, in
the \Rpackage{iSNetwork} package there will be the concept of the ``active''
MVC object, where the ``active'' MVC object (and thus, the ``active'' model)
means that operations performed through the control window menus or through
the command line functions will affect that MVC object.  For
example, the Display menu allows users to create views of the active model,
but only certain views are appropriate for a type of model, as mentioned in
Section \ref{Ssec:OneViews}.  If the active model was a data frame model
(\Robject{dfModel}), then the views that the user could create would be a
scatterplot (\Robject{sPlotView}) and a spreadsheet (\Robject{spreadView}),
but a graph plot (\Robject{graphView}) would not make sense.  Thus, when the
model in the ``active'' MVC is of class \Robject{dfModel}, then the Display
menu will have menu items to create a scatterplot and a spreadsheet.  Also, if
the ``active'' MVC is of class \Robject{dfModel}, then the command line
functions will only let the user create a scatterplot and a spreadsheet view
of the active model.

Because one of the goals of creating linked views based on multiple linked
data sets is to create a software package that is flexible and extensible,
users will be able to add menus and menu items to the control window through
the command line to access new functionality that they have created.  As an
example if a future programmer wants to add a new type of view, such as a
histogram, then a new menu item for that view could be added to the Display
menu, as long as that view makes sense for the active MVC's model.  By
allowing users to change the controller, they can add new functionality that
is specific to their problem.  Note that when users make additions to the API,
they should add new command line functions in addition to adding menus and
menu items to access that functionality.

\subsubsection{Linking Events to Functions}\label{Ssec:OneEvent}

Linking callback functions to events, such as a button click, a mouse movement
or a key press event, is what allows users to have an interactive
environment.  When an event occurs, the callback function linked to the event
will determine the response to that event.  To create a flexible environment
it would be nice to be able to change the response, which is determined by the
callback function, to an event.  For example, when a left button press event
occurs, a user may sometimes want a point to be colored and at other times she
may want a point to be hidden.

To allow this flexibility in the response to an event, a class called
\Robject{gEventFun} was created.  The \Robject{gEventFun} class stores all of
the information needed to create a particular response to an event.
Clearly, the response to an event can be performed by a single function,
which is called the callback function, as discussed in Section
\ref{Ssec:Backg}.  However, different responses to an event may require
functions with different signatures.  Using the example in the previous
paragraph, a callback function that colored a point would need two parameters,
the new color to use and the point to color, while a callback function that
hid a point would need just one parameter, the point to hide.  Since the
intention is to be able to interchange the responses to an event, the
signatures of different callback functions should be the same.  These design
decisions determined the slots in the \Robject{gEventFun} class and will be
discussed more fully in the following paragraphs.

As shown in Figure \ref{Fig:EventFun}, the slots in the \Robject{gEventFun}
class are \Rslot{callFun}, \Rslot{shortName}, and \Rslot{preprocessFun}.  The
\Rslot{callFun} slot is the name of the callback function, the
\Rslot{shortName} slot is a short description of the function, and the
\Rslot{preprocessFun} slot is a character vector of all preprocessing
functions that must be called before the callback function.

\begin{figure}[ht]
  \begin{center}
    \scalebox{0.6}{\includegraphics{EventClass.jpg}}
    \caption{ The gEventFun Class. }
    \label{Fig:EventFun}
  \end{center}
\end{figure}

The callback function that is stored in the \Rslot{callFun} slot must take
only one parameter.  As discussed previously, all callback functions must have
the same signature so that different callback functions can be interchanged in
the code to allow different responses to an event.  All callback functions
will need to know what object they should act on so this will be the one
parameter that the callback functions take.  Any other information a callback
function may need will be set using the preprocessing functions.  The class of
the one parameter for the callback function will depend on the model class in
the active MVC object.  For example, for a graph model, the parameter will be
of class \Robject{AgNode}, representing a node in the graph.  For a data frame
model, the parameter will be of class character, representing a row in the
data frame.

The preprocessing functions that are stored in the preprocessFun slot will
take no parameters.  The purpose of the preprocessing functions is to set some
variables that the callback function will need.  Thus, not all callback
functions will need preprocessing functions and if no preprocessing functions
are needed, then this slot will be NULL.  If the callback function sets the
color of a node in a graph, then a preprocessing function is needed to
determine what color this will be.  So in this example, the preprocessing
function will open a color browser to allow the user to choose what the new
color will be.

When a user wants to link a callback function stored in an \Robject{gEventFun}
instance to an event, this can be performed either through the GUI or through
the command line functions.  The controller environment in each MVC object,
which will be discussed in detail in Section \ref{Ssec:OneMVC},
will store which \Robject{gEventFun} object (representing the callback
function) is linked with each event.  Note that not all callback functions
make sense for each event so the controller environment will also store a list
for each event of appropriate \Robject{gEventFun} instances that can be linked
to that event.

\subsection{MVC}\label{Ssec:OneMVC}

The \Robject{MVC} class is the class that binds the model, view, and
controller objects that refer to the same data set into one object.  Thus, the
creation of a MVC object occurs whenever a new data set is loaded.  A
MVC object can have one model and each model must be in one
MVC object so there is a one-to-one relationship between the
MVC and model objects.  Because of this relationship, the
MVC object and the model object are identified by the same name,
which is stored in the \Rslot{modelName} slot of the model object.  When
loading a new model through \Rpackage{iSNetwork's} API, the name of the model
will be checked to ensure that it is unique.

As shown in Figure \ref{Fig:MVCClass}, the \Robject{MVC} class has five slots
which are, \Rslot{model}, \Rslot{viewList}, \Rslot{controller},
\Rslot{parentMVC}, and \Rslot{childMVCList}.  The \Rslot{model} slot stores
the model object, the \Rslot{viewList} slot is a list of the view objects for
this model, the \Rslot{controller} slot is an environment that stores
information for this MVC, the \Rslot{parentMVC} slot is the name of the parent
model (MVC), and the \Rslot{childMVCList} slot is a list of the children models
(MVCs).  Note that model and MVC objects use the same name so the
\Rslot{parentMVC} slot refers to both the name of the model and the name of the
MVC.  Similarly, for the \Rslot{childMVCList} slot, the names in the list
refer to both the name of the model and the name of the MVC.

\begin{figure}[ht]
  \begin{center}
    \scalebox{0.6}{\includegraphics{MVCClass.jpg}}
    \caption{ The MVC Class. }
    \label{Fig:MVCClass}
  \end{center}
\end{figure}

Thus, the \Robject{MVC} class not only binds the model, view, and controller
objects together, but it also shows that the MVC objects are related
to each other through a parent-child (tree) hierarchy, as shown in Figure
\ref{Fig:Hier}.  In this picture, `model1' is the name of the original MVC
object (model) and from it, two MVC objects were created, `model2' and
`model3'. Then from `model2', three MVC objects were created, `model4',
`model5', and `model6'.

%\clearpage

\begin{figure}[ht]
  \begin{center}
    \scalebox{0.6}{\includegraphics{Hierarchy2.jpg}}
    \caption{ The Parent-Child Relationship Between MVCs. }
    \label{Fig:Hier}
  \end{center}
\end{figure}

A design decision was that a new model (and thus, a new MVC object) could
either be loaded or could be created from an already loaded MVC object.  As
mentioned in Section \ref{Ssec:MMVC}, the focus in this paper will be on
experimental data that are linked to meta-data, as the linked data sets that we
are most interested in visualizing.  In this situation, the experimental data
(and its MVC object) can be thought of as determining what meta-data are
studied.  Thus, there is a hierarchical relationship between these data sets
where one MVC object creates another MVC object.  This focus led to the
decision that a MVC object can have at most one parent MVC, from which it was
derived, but it can have many child MVCs.  An easy example of a child model
derived from a parent model is to start with a data frame as the parent model
and then take a subset of this data frame as the child model.

\subsection{Messages Within a MVC Object}\label{Ssec:OneMess}

The message classes are intended to provide communication between different
components of the MVC design, such as when the model changes a message needs to
be sent to the views to let them know that they should be updated.  This
communication between the model, view and controller is crucial for the pieces
to work together and yet, still be independent of each other.  For information
about messages that are passed between MVC objects, please see Section
\ref{Ssec:MultMess}.

The object model for the message classes was derived from the idea that all
messages must have certain methods in common, such as \Rfunction{initialize}
and \Rfunction{handleMessage}, because all messages must be created and
handled in some manner so that the message is read and acted upon.  The
\Rfunction{handleMessage} method can be considered a central dispatch for
messages because all messages will be sent to the \Rfunction{handleMessage}
method, which will determine where the information should be sent.   Also,
since messages are sent when something has changed, either through addition,
alteration, or deletion, the message classes reflect only certain operations.
These common purposes for the messages allow the message object model to be
constructed, and this model is shown in Figure \ref{Fig:Mess}.

\begin{figure}[ht]
  \begin{center}
    \scalebox{0.6}{\includegraphics{MessageClass.jpg}}
    \caption{ Inheritance for Message Objects. }
    \label{Fig:Mess}
  \end{center}
\end{figure}

As with the view classes, the top message class, \Robject{gMessage}, is a
virtual class.  It contains no slots and its purpose is to bind the other
message classes together.

Currently, there are two types of messages for messaging within a model, view,
controller object: an add and an update message.  Both the add message,
\Robject{gAddMessage}, and the update message, \Robject{gUpdateMessage},
contain three slots, \Rslot{dataName}, \Rslot{mData}, and \Rslot{type}.
\Rslot{dataName} is the character string that gives the name of the model,
\Rslot{mData} is a list of data needed to perform the addition or alteration
operation, and \Rslot{type} is a character string that gives the type of
addition or alteration to perform.  Both \Robject{gAddMessage} and
\Robject{gUpdateMessage} are virtual classes.

Since adding a view requires different information and methods from adding a
model (and similarly with updating views versus models), it makes sense to
have two separate message classes for these operations.  Thus, when messages
are sent between components, they are from one of the following classes:
\Robject{gAddDataMessage}, \Robject{gAddViewMessage},
\Robject{gUpdateDataMessage}, and \Robject{gUpdateViewMessage}.  As shown in
Figure \ref{Fig:Mess}, the \Robject{gAddDataMessage} and
\Robject{gAddViewMessage} classes inherit from the \Robject{gAddMessage}
class, and the \Robject{gUpdateDataMessage} and \Robject{gUpdateViewMessage}
classes inherit from the \Robject{gUpdateMessage} class.

The \Robject{gAddDataMessage} class represents a message to add a model and a
MVC object because whenever a model is added, a MVC object is also added that
contains this model.  The \Robject{gAddViewMessage} class represents a message
to create a view.  This message can only be used after a model has been
added.

The \Robject{gUpdateDataMessage} class represents a message to update
a model.  This message is crucial for linking because it is responsible for
ensuring that once a model has been updated this information propagates to
any views of the model.  This same information must also be passed along to any
other MVC object that is related to this model.  Because this information gets
passed to other MVC objects, which will be discussed further in Section
\ref{Sec:MultMVC}, this message also has one new slot, \Rslot{from}, in
addition to the slots it inherits from \Robject{gUpdateMessage}.  The
\Rslot{from} slot is the name of the model that the update data message came
from.  This slot is necessary because the update data message can come from
one of two sources: the message can occur because of user interaction with a
view of this model or the message may occur because a different model that is
linked to this model has been updated.  So the update data message may come
from within this MVC object or it may come from a different MVC object that
has changed its model and is related to this MVC object.  Information on
passing messages between MVC objects will be given in Section
\ref{Sec:MultMVC}.  Finally, the \Robject{gUpdateViewMessage} class represents
a message to update a view.  This message will be created and handled whenever
the model is updated so that the views accurately reflect the model.

All four classes, \Robject{gAddDataMessage}, \Robject{gAddViewMessage},
\Robject{gUpdateDataMessage}, and \Robject{gUpdateViewMessage}, have
\Rfunction{initialize} and \Rfunction{handleMessage} methods, which perform
the following tasks:

\begin{list}{}{\setlength{\leftmargin}{0.5in}
               \setlength{\itemsep}{0in}}
  \item \Rfunction{initialize} is used to properly fill the slots of the
object when it is created.  
  \item \Rfunction{handleMessage} is necessary to
process the message.  Since a message is a form of communication
between two components, once it is created it must be sent to the
proper recipient and action must be taken in response to its
information, which is the purpose of the \Rfunction{handleMessage} method.
\end{list}

As with the inheritance structures seen in previous sections, the message
inheritance structure is designed to allow for extensibility.  For example, a
user could create a \Robject{gUpdateControlMessage} class that inherited from
the \Robject{gUpdateMessage} class.  This new class,
\Robject{gUpdateControlMessage}, could convey information that told the
control window to be updated in response to a change in the active view (the
active view is the view that has the focus).  Currently, the control window
does not reflect what the active view is, but if a view was created that could
be rotated, then a user may want the control window to have some type of
interface to allow the user to control the rotation.  The goal for the message
classes is that the design is flexible enough to allow a user to make
additions or changes.

\section{Details of Multiple MVCs}\label{Sec:MultMVC}

We now describe the implementation of multiple MVC objects for linking views
of multiple linked data sets.  Each MVC object will represent one model and
its views.  Now for the views of different data sets to be linked, they must
link through the data sets so this linking information must be stored
somewhere.  Also, there must be message classes to pass information between
MVC objects.

Thinking of experimental data that are linked to meta-data as an example, there
will be two MVC objects, one for the experimental data and one for the
meta-data.  In the example of experimental microarray data that are linked to
the Gene Ontology molecular function graph that was first mentioned in Section
\ref{Ssec:Limit}, these two data sets are linked in the following manner: the
Affymetrix identifiers in the experimental microarray data are linked to Locus
Link identifiers, which are in turn linked to the Gene Ontology molecular
function terms.  Thus, to pass information between these two data sets a table
that links Affymetrix identifiers to Gene Ontology molecular function terms
must be stored or there must be a function to call that converts from
Affymetrix identifiers to Gene Ontology molecular function terms and vice
versa.  Once it is possible to link these data sets, then information must be
passed between them through messages to keep the views synchronized.

As mentioned in Section \ref{Ssec:OneMVC}, the \Robject{MVC} class has slots
for how the current MVC relates to other MVC objects and these slots are
\Rslot{parentMVC} and \Rslot{childMVCList}.  These slots show that MVC objects
are related through a tree hierarchy where a child model is derived from its
parent model.  Consider Figure \ref{Fig:SmallHier} shown below.  If firstMVC's
model changes, then this information first gets passed to secondMVC, which may
make changes to its model based on this information.  Then secondMVC passes
this information to thirdMVC based on the changes secondMVC made to its model.
Thus, information passing between firstMVC and thirdMVC only occurs through
secondMVC.  So messages can only be sent from parent MVC to child MVC or from
child MVC to parent MVC.

\begin{figure}[ht]
  \begin{center}
    \scalebox{0.6}{\includegraphics{SmallHierarchy.jpg}}
    \caption{ Multiple MVC Objects and Their Relationships. }
    \label{Fig:SmallHier}
  \end{center}
\end{figure}

Next we will consider how the data sets will be linked (and thus, how MVC
objects are linked) followed by how messages will be passed between these
linked MVC objects.

\subsection{Linking Data Sets}\label{Ssec:MultLink}

For messages to pass between MVC objects, there must be a function or a place
that converts information from one MVC into useful information for a different
MVC object.  Thus, there needs to be some way to determine how data from one
MVC object relates to data in a different MVC object.  This conversion is done
by two functions that are stored in the \Rslot{linkData} slot of the model
objects.  The \Rslot{linkData} slot is a list with two elements: a
\Rfunction{toParent} function and a \Rfunction{fromParent} function.

When a child MVC is created from an existing MVC, the child model's
\Rslot{linkData} slot will be filled with the two functions,
\Rfunction{toParent} and \Rfunction{fromParent}.  When the model in the child
MVC is updated, the \Rfunction{toParent} function in the \Rslot{linkData} slot
will be used to convert the child model's data change into useful information
for the parent model.

Now suppose that the parent model changed instead.  This information
needs to be passed to the child model.  However, a parent can have multiple
child MVC objects so the parent must look in each child model's
\Rslot{linkData} slot for the \Rfunction{fromParent} function.  The
\Rfunction{fromParent} function will convert the parent model's data change
into useful information for this particular child model.  Thus, the parent
model's data change may be converted into different information for different
child models and this can happen because each child model has its own
\Rfunction{fromParent} function.  When a child MVC object is created and
its model's \Rslot{linkData} slot is filled, the \Rfunction{toParent} function
will convert information from this child model into useful information for its
parent MVC and the \Rfunction{fromParent} function will convert a message
from its parent MVC into useful information for this child MVC.

An example of how these functions, \Rfunction{toParent} and
\Rfunction{fromParent}, are used when messages are passed between MVCs will
be given in Section \ref{Ssec:MultMess}.

\subsection{Messages Between MVC Objects}\label{Ssec:MultMess}

In Section \ref{Ssec:OneMess}, where the message classes to pass information
within one MVC object were discussed, it was noted that there are two main
types of messages: an add message and an update message.  These message types
will also be seen in passing messages between different MVC objects, as shown
in Figure \ref{Fig:BetwMess}.

\begin{figure}[ht]
  \begin{center}
    \scalebox{0.6}{\includegraphics{newMessageClass2.jpg}}
    \caption{ Inheritance for Message Classes that are Passed Between MVCs. }
    \label{Fig:BetwMess}
  \end{center}
\end{figure}

The first message class that will be discussed is \Robject{gAddChildMessage}.
A \Robject{gAddChildMessage} object will be very similar to a
\Robject{gAddDataMessage} object because they are both adding a new
model and a new MVC object.  The difference with a \Robject{gAddChildMessage}
object is that it must fill in some extra information to tie this new MVC
object to its parent MVC object.  The \Robject{gAddChildMessage} class
inherits from the \Robject{gAddMessage} class, with the same slots of
\Rslot{dataName}, \Rslot{mData}, and \Rslot{type}.  The \Rslot{dataName} slot
is the name of the new model (and new MVC), the \Rslot{mData} slot is a list
that contains the model data and virtual data to fill the \Rslot{modelData}
and \Rslot{virtualData} slots, respectively, of the new model, and the
\Rslot{type} slot gives the type of the model (currently, the options are
``exprSet'', ``graph'', or ``data.frame'').  As with the previous message
classes, the \Robject{gAddChildMessage} class has two methods:
\Rfunction{initialize} and \Rfunction{handleMessage}.  The
\Rfunction{initialize} method properly fills the slots of the object when it
is created and the \Rfunction{handleMessage} method ensures that the new model
and MVC objects are created (which is what the \Rfunction{handleMessage}
method for the \Robject{gAddDataMessage} class does) and then it fills the
\Rslot{linkData} slot of the new child model so that the \Rfunction{toParent}
and \Rfunction{fromParent} functions are available for messaging between MVC
objects and finally it stores information in the child MVC object's controller
that determines if the child model was created from a subset of the parent
model.  Thus, the \Rfunction{handleMessage} method for the
\Robject{gAddChildMessage} class both creates the child MVC object and relates
it to the parent MVC object.

The next two message classes pertain to sending information to linked MVC
objects when a parent MVC or child MVC object has changed.  These classes are
\Robject{gSendParentMessage} and \Robject{gSendChildMessage} and they both
inherit from the \Robject{gMessage} class, as shown in Figure
\ref{Fig:BetwMess}.  Because the messages are sending information to a linked
MVC object, they both contain a slot for a \Robject{gUpdateDataMessage} object
that was just used to update a MVC object.  The \Robject{gSendParentMessage}
class has one slot, \Rslot{childUpdateDataMessage}, and this slot contains the
\Robject{gUpdateDataMessage} object that was used to update the child model.
Similarly, the \Robject{gSendChildMessage} class has two slots,
\Rslot{parentUpdateDataMessage} and \Rslot{childName}, where the
\Rslot{parentUpdateDataMessage} slot contains the \Robject{gUpdateDataMessage}
object that was used to update the parent model and the \Rslot{childName} slot
contains the name of the child model that is being updated because a parent
MVC can have more than one child MVC.

As with all the other message classes, these two send messages have two
methods: \Rfunction{initialize} and \Rfunction{handleMessage}.  The
\Rfunction{initialize} method properly sets the slots of the new object.  The
\Rfunction{handleMessage} method for a \Robject{gSendParentMessage} will
take the \Robject{gUpdateDataMessage} object from the child model and convert
it to a \Robject{gUpdateDataMessage} object for the parent model using the
\Rfunction{toParent} function.  Similarly, the \Rfunction{handleMessage}
method for a \Robject{gSendChildMessage} will take the
\Robject{gUpdateDataMessage} object from the parent model and convert it to a
\Robject{gUpdateDataMessage} object for the child model using the
\Rfunction{fromParent} function.  An example of how these messages are passed
and converted is shown in Figure \ref{Fig:MessPass}.

These message classes ensure that when a model is updated, its parent and
child models will be notified of the change and most importantly will be
notified of the change in a way that each model can properly act upon that
information.  Because each MVC object has a different model (and thus a
different data set and potentially a different data structure), information
that pertains to the model for one MVC must be converted to useful information
for the linked model and this conversion is done by the \Rfunction{toParent}
and \Rfunction{fromParent} functions.

\begin{figure}[ht]
  \begin{center}
    \scalebox{0.55}{\includegraphics{newMessagePassing.jpg}}
    \caption{ Message Passing Between and Within MVCs. }
    \label{Fig:MessPass}
  \end{center}
\end{figure}

Figure \ref{Fig:MessPass} is very similar to Figure \ref{Fig:firstMP} shown in
Section \ref{Ssec:Backg}.  Now, however, Figure \ref{Fig:MessPass} shows how
MVC objects are defined and how messages are passed within one MVC as well
as how this information gets passed on to its parent MVC using messages.  In
this picture, the message passing starts when a user interacts with the y
vs. x scatterplot depicting model 2.  After the user interaction with the
scatterplot, step 1 (indicated on the figure as a circled 1) is to send an
update data message to model 2.  As soon as model 2 is updated, step 2 is to
send an update view message so that the scatterplot view is updated to reflect
the changed data.  In step 3, a send parent message is sent because MVC 2
(which is the MVC object for model 2) has a parent MVC, which is MVC 1.  For
step 4, the \Rfunction{toParent} function converts the child's update data
message into an update data message that the parent will understand.  In step
5, this new update data message that was created by the \Rfunction{toParent}
function will be sent to model 1.  Finally for step 6, after model 1 has been
updated, an update view message will be sent to the view of model 1.  Now all
of the data and the views are synchronized.

\section{Conclusions}\label{Sec:Conc}

Several examples of linked data sets were given at the beginning of this
paper, including linked tables from a relational database, and meta data that
are linked to experimental data.  The model-view-controller (MVC) paradigm was
discussed as a useful design for creating linked views of a data set.
Starting with the MVC paradigm, this paper provided a design for implementing
linked, interactive views of linked data sets based on expanding the MVC
paradigm.  For visualizing linked data sets, the MVC design, which is a well
known and thus, well tested software paradigm, was extended to multiple MVC
objects where each MVC object contains one data set and its views.  This
solution broke down the problem of linked, interactive views of linked data
sets into components (MVC objects) that can be reused, extended and linked.

Once the design focused on MVC objects as the component of interest for
creating linked views of linked data sets, the next step was to determine how
these MVC objects would be related.  Because the focus in this paper was on
experimental data that are linked to meta-data, a hierarchical relationship
between MVC objects was built into the MVC class definition.  The experimental
data can be thought of as determining what meta-data are studied and thus, the
meta-data are derived from the experimental data, which shows a parent-child
relationship.

Another consideration in the design was creating a method for communicating
between MVC objects.  Already communication within a MVC object was sent using
instances of the message classes.  Thus, it seemed a natural extension to have
message classes that pertained to passing information between MVC objects.
The goal of the message classes was to allow communication between independent
components in the design.

The goals for this paper were to create linked views of linked data sets, to
create interactive views where the response to an event can be changed, and to
create an extensible design that will allow future users to make additions.
The first goal was accomplished by implementing MVC objects as components that
communicate with each other through messages.  The second goal of creating
interactive views was met by using the \Rpackage{RGtk} and
\Rpackage{gtkDevice} packages to create devices that could respond to events.
In addition the \Robject{gEventFun} class was designed so that users could
alter the response to an event.  Finally, the last goal of creating an
extensible design was achieved through the inheritance structure for classes,
which allows for extensions by the user, and through command line functions,
which let the user create new menu items to add to the GUI and also let the
user create new callback functions that can be linked to events.

Although extensibility was a goal, certain design decisions, such as expecting
the data sets to be linked through 
a parent-child hierarchy, may not work with some data sets.  As an example,
the linked tables from a relational database are not related through a tree
structure and thus, having parent and child MVC objects may not make sense
with these data sets.  However, the design allows users to add new
types of models, new types of views, and new message classes so there is still
flexibility for additions as long as the data sets are connected through
parent-child relationships.  These new class definitions could be created by
the user in a new package that extends the \Rpackage{MVCClass} package that
was discussed in this paper.  Other additions that the user may want to
perform, such as adding new menus to the GUI and adding new responses to an
event can be performed through command line functions that are defined in the
\Rpackage{iSNetwork} package.

\subsection{Future Work}\label{Ssec:Further}

When a MVC's model is updated, it notifies its parent and child
MVCs that a change has occurred.  Then when a MVC gets a message from a
parent or child MVC, the MVC can either decide to ignore the message or read
the message, which results in this MVC's model being updated.  However, the
MVC object can not decide to only see certain types of messages.  It must
either accept all messages and update its model based on the messages, or
ignore all messages.  A future goal for message passing is to allow more
selectivity in which messages will be accepted between MVC objects so that the
linking between MVC objects is more sophisticated.

\end{document}