### 2006-07-20 inst/doc/MVCClass.Rnw
Changed the text for the vignette, added a man page, and removed some old pictures that are not being used.
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MVCClass@18900 bc3139a8-67e5-0310-9ffc-ced21a209358

%
% NOTE -- ONLY EDIT THE .Rnw FILE!!!  The .tex file is
% likely to be overwritten.
%
%\VignetteDepends{RGtk, gtkDevice, graph, Rgraphviz}
%\VignetteKeywords{Model, View, Controller}
%\VignettePackage{MVCClass}
%\VignetteIndexEntry{MVCClass}

\documentclass[11pt]{article}

\usepackage{amsmath,pstricks}
\usepackage[authoryear,round]{natbib}
\usepackage{hyperref}
\usepackage{times}
\usepackage{comment}
\usepackage{epsfig}
\usepackage{graphics}
\usepackage{graphicx}

\textwidth=6.2in
\textheight=8.5in
%\parskip=.3cm
\oddsidemargin=.1in
\evensidemargin=.1in
\headheight=-.3in

\newcommand{\scscst}{\scriptscriptstyle}
\newcommand{\scst}{\scriptstyle}

\newcommand{\Rfunction}[1]{{\texttt{#1}}}
\newcommand{\Robject}[1]{{\texttt{#1}}}
\newcommand{\Rpackage}[1]{{\textit{#1}}}

\bibliographystyle{plainnat}

\title{Model-View-Controller Package}
\author{Elizabeth Whalen}
\date{July 19, 2006}

\begin{document}
\maketitle

\section{Overview}\label{Sec:Overview}

The purpose of the \Rpackage{MVCClass} package is to provide the definitions of
classes and generic functions that will be used by other packages to create 
model-view-controller applications.  Currently, the \Rpackage{iSPlot} and the
\Rpackage{iSNetwork} packages use the \Rpackage{MVCClass} package to create
linked, interactive views of data.  The \Rpackage{BioMVCClass} package extends
the classes in the \Rpackage{MVCClass} package to include more model and view
classes.  These definitions were put in a separate R package because the
\Rpackage{BioMVCClass} packages depends on the following R packages:
\Rpackage{Biobase}, \Rpackage{graph}, and \Rpackage{Rgraphviz}.

\section{MVC Class}\label{Sec:MVC}

The \Robject{MVC} class represents a model-view-controller object.  Basically, 
this class will bind the model with its views and its controller so that 
an object of this class can be created each time there is a new data set 
(model).  Thus, there is a one-to-one relationship between a MVC object and
a model object.  Because of this relationship, the
MVC object and the model object are identified by the same name,
which is stored in the modelName slot of the model object (see Section 
\ref{Sec:Model}).  Figure \ref{Fig:MVCClass} 
shows the inheritance structure for the \Robject{MVC} classes.

\begin{figure}[ht]
  \begin{center}
    \scalebox{0.2}{\includegraphics{fullMVCClass.jpg}}
    \caption{ Inheritance for MVC Objects. }
    \label{Fig:MVCClass}
  \end{center}
\end{figure}

The \Robject{MVC} class is a virtual class that binds the other \Robject{MVC}
classes together.  Every \Robject{MVC} object will need the three slots:
model, viewList and controller.  The model slot stores the model object, the
viewList slot is a list of the view objects for this model, and the controller
slot is an environment that stores information for this MVC.  Inheriting from
the \Robject{MVC} class is the \Robject{singleModelMVC} class that has no new
slots.  The \Robject{singleModelMVC} class represents a \Robject{MVC} object
that is not linked to any other \Robject{MVC} objects.

Inheriting from the \Robject{singleModelMVC} class is the
\Robject{linkedModelMVC} that contains the new slots: parentMVC and
childMVCList.  The parentMVC slot is the name of the parent model (MVC) and
the childMVCList slot is a list of the children models (MVCs).  Note that
model and MVC objects use the same name so the parentMVC slot refers to both
the name of the model and the name of the MVC.  Similarly, for the
childMVCList slot, the names in the list refer to both the name of the model
and the name of the MVC.

Thus, the \Robject{singleModelMVC} class binds the model, view, and controller
objects together and the \Robject{linkedModelMVC} class allows the MVC objects
to be related to each through a parent-child (tree) hierarchy because of the
slots, parentMVC and childMVCList.  These two slots also show that a MVC
object can have at most one parent MVC and can have many children MVCs.

\section{Model Classes}\label{Sec:Model}

The model class is responsible for storing and updating the data.  These two
functions are reflected in Figure \ref{Fig:Model}, which shows the inheritance
structure for the model classes.  Here the virtual class,
\Robject{gModel}, has the slots: modelData, linkData, virtualData, modelName,
and modelVar.  These five slots are common information that all models need.
The modelData slot is the data set for this model, the linkData slot is
a list of two functions, \Rfunction{toParent} and \Rfunction{fromParent},
which link this model to its parent and child models, respectively (see Section
\ref{Sec:MVC}), the virtualData slot is data pertaining to the views
that needs to be stored with the model so that it can be shown in all views of
this model, the modelName slot is the name of the model (and the name of the
MVC), and the modelVar slot is a list of named model variables (for example, a
statistic for each element in the model).

\begin{figure}[ht]
  \begin{center}
    \scalebox{0.7}{\includegraphics{VigModelClass.jpg}}
    \caption{ Inheritance for Model Objects. }
    \label{Fig:Model}
  \end{center}
\end{figure}

In this package, the only specific model class is \Robject{dfModel}.
The \Robject{dfModel} class represents a model where the modelData has class
data frame (or matrix).  Other model classes are defined in the
\Rpackage{BioMVCClass} package.

As mentioned previously, the model classes are also responsible for updating
the data so a generic function \Rfunction{updateModel} has also been 
defined.  The \Rfunction{updateModel} methods will be defined in packages that 
use this package (i.e. the methods will be defined in \Rpackage{iSNetwork} and 
\Rpackage{iSPlot}).

The model classes may also be asked to provide information by a
\Robject{gAskAncestorMessage} object (which is discussed in Section
\ref{Sec:Message}) and thus, the generic function \Rfunction{provideInfo} has
been defined.  Again, the \Rfunction{provideInfo} methods are defined in
packages that use this package, such as \Rpackage{iSNetwork}.

\section{View Classes}\label{Sec:View}

The view classes represent the visual depictions of the model.  All views will
need to store some common information and respond to certain events through
methods.  This consideration led to an object model where the different view
classes inherit from a view virtual class, called \Robject{genView}.  The
object model for the view classes is shown in Figure \ref{Fig:View}.

\begin{figure}[ht]
  \begin{center}
    \scalebox{0.7}{\includegraphics{VigViewClass.jpg}}
    \caption{ Inheritance for View Objects. }
    \label{Fig:View}
  \end{center}
\end{figure}

The \Robject{genView} class has three slots, dataName, win, and winNum.
dataName is the name of the model that the view displays, win is the Gtk
window object that holds the view, and winNum is the number of the window so
that the window can be identified in the GUI that other packages create.  All 
views will need to know these three pieces of information so \Robject{genView} 
will bind the view classes together.

As for specific view classes, the \Robject{spreadView} class will represent a
spreadsheet view of the data.  This view will only make sense for models that
are a two dimensional data structure, such as a matrix or a data frame.  The
information needed for this view is stored in the slot, clist, which is the
Gtk spreadsheet object.

When creating a plot of a data set, there is a general class, called
\Robject{plotView}, that has the slots, plotDevice, plotPar, and drArea, which
store the device number of the plot, the plotting parameters, and the Gtk
drawing area object, respectively.  This class is a virtual class because it
is not intended to have any objects as it just represents a general plot.  The
plot classes that can have objects are the specific plot classes,
\Robject{sPlotView} and \Robject{qqPlotView}, which inherit from the
\Robject{plotView} class.  More view classes are defined in the
\Rpackage{BioMVCClass} package.

The \Robject{sPlotView} class represents a scatterplot view and it has the
extra slots, dfRows, xvar, and yvar, where dfRows are the row names from the
model that are shown in the plot, xvar is the variable name (the column name
if the model is a data frame) from the model that
is shown as the x variable in the plot, and yvar is the variable name (the
column name if the model is a data frame) from the
model that is shown as the y variable in the plot.  The \Robject{qqPlotView}
represents a qq-plot view and it has the extra slots, xval and yval, where
xval are the numeric values of the points on the x-axis and yval are the
numeric values of the points on the y-axis.

Several generic functions have been defined that refer to operations 
performed on views.  These generic functions are \Rfunction{motionEvent}, 
\Rfunction{clickEvent}, \Rfunction{updateView}, \Rfunction{redrawView} and
\Rfunction{identifyView}.  Methods for these generic functions are defined in
packages that use this package, such as \Rpackage{iSPlot} and
\Rpackage{iSNetwork}.  When these methods are defined, the
\Rfunction{motionEvent} method responds to a mouse over event on a view, the
\Rfunction{clickEvent} method responds to a click event over a view, the
\Rfunction{updateView} method updates only a portion of the view, the
\Rfunction{redrawView} method completely redraws the view, and the
\Rfunction{identifyView} method identifies an object (such as a point, a node,
etc.) on a view.

\section{Message Classes}\label{Sec:Message}

The message classes are intended to provide communication between different
components of the MVC design, such as when the model changes a message needs to
be sent to the views to let them know that they should be updated.  This
communication between the model, view and controller is crucial for the pieces
to work together and yet, still be independent of each other.

The object model for the message classes was derived from the idea that all
messages must have certain methods in common, such as \Rfunction{initialize}
and \Rfunction{handleMessage}, because all messages must be created and
handled in some manner so that the message is read and acted upon.  Also,
since messages are sent when something has changed, either through addition,
alteration, or deletion, the message classes reflect only certain operations.
These common purposes for the messages allow the message object model to be
constructed, and this model is shown in Figure \ref{Fig:Mess}.

\begin{figure}[ht]
  \begin{center}
    \scalebox{0.7}{\includegraphics{VigMessageClass.jpg}}
    \caption{ Inheritance for Message Objects. }
    \label{Fig:Mess}
  \end{center}
\end{figure}

The figure shows the inheritance structure for message objects that can 
be passed within one MVC object and Figure \ref{Fig:BetwMess} in this section
shows the inheritance structure for messages objects that can be passed
between MVC objects.

As with the view classes, the top message class, \Robject{gMessage}, is a
virtual class.  It contains no slots and its purpose is to bind the other
message classes together.

Currently, there are two types of messages for messaging within a model, view,
controller object: an add and an update message, and both of these messages
are performing some type of modification.  Thus, both the add message,
\Robject{gAddMessage}, and the update message, \Robject{gUpdateMessage},
inherit from the virtual class, \Robject{gModifyMessage}, which has 
three slots, dataName, mData, and type.  dataName is the character
string that gives the name of the model, mData is a list of data needed to
perform the addition or alteration operation, and type is a character string
that gives the type of addition or alteration to perform.  Both
\Robject{gAddMessage} and \Robject{gUpdateMessage} are also virtual classes.

Since adding a view requires different information and methods from adding a
model (and similarly with updating views versus models), it makes sense to
have two separate message classes for these elements.  Thus, when messages
are sent between components, they are from one of the following classes:
\Robject{gAddDataMessage}, \Robject{gAddViewMessage},
\Robject{gUpdateDataMessage}, and \Robject{gUpdateViewMessage}.  As shown in
Figure \ref{Fig:Mess}, the \Robject{gAddDataMessage} and
\Robject{gAddViewMessage} classes inherit from the \Robject{gAddMessage}
class, and the \Robject{gUpdateDataMessage} and \Robject{gUpdateViewMessage}
classes inherit from the \Robject{gUpdateMessage} class.

The \Robject{gAddDataMessage} class represents a message to add a model and a
MVC object because whenever a model is added, a MVC object is also added that
contains this model.  The \Robject{gAddViewMessage} class represents a message
to create a view.

The \Robject{gUpdateDataMessage} class represents a message to update
a model.  This message is crucial for linking because it is responsible for
ensuring that once a model has been updated this information must propagate to
any views of the model and this information must also be passed along to any
other MVC object that is related to this model.  Because this information gets
passed to other MVC objects, this message also has one new slot, from, in 
addition to the slots it inherits from \Robject{gUpdateMessage}.  The from
slot is the name of the model that the update data message came from.  This
slot is necessary because the update data message can come from one of two
sources: the message can occur because of user interaction with a view
of this model or the message may occur because a different model that is
linked to this model has been updated.  So the update data message may come
from within this MVC object or it may come from a different MVC object that
has changed its model and is related to this MVC object.  Information on
passing messages between MVC objects is given later in this section.  Finally,
the \Robject{gUpdateViewMessage} class represents a message to update a view.

Message classes have also been defined to pass information between MVC
objects.  The message types for passing information between \Robject{MVCs} are
similar to the messages that are passed within \Robject{MVC} objects.  Now the
add child message is similar to the add model message, and the send parent
message and the send child message are similar to the update model message.
The inheritance structure for messages passed between MVC objects is shown in
the Figure \ref{Fig:BetwMess}.

%\clearpage

\begin{figure}[ht]
  \begin{center}
    \scalebox{0.7}{\includegraphics{newVigMessageClass2.jpg}}
    \caption{ Inheritance for Message Classes that are Passed Between MVCs. }
    \label{Fig:BetwMess}
  \end{center}
\end{figure}

The first message class that will be discussed is \Robject{gAddChildMessage}.
A \Robject{gAddChildMessage} object is very similar to a
\Robject{gAddDataMessage} object because they are both adding a new
model and a new MVC object.  The difference with a \Robject{gAddChildMessage}
object is that it must fill in some extra information to tie this new MVC
object to its parent MVC object.  The \Robject{gAddChildMessage} class
inherits from the \Robject{gAddMessage} class, with the same slots of
dataName, mData, and type.  The dataName slot is the name of the new model (and
new MVC), the mData slot is a list that contains the model data and virtual
data to fill the modelData and virtualData slots, respectively, of the new
model, and the type slot gives the type of the model.

The next two message classes pertain to updating MVC objects when a parent MVC
or child MVC object has changed.  These classes are
\Robject{gSendParentMessage} and \Robject{gSendChildMessage} and they both
inherit from the \Robject{gMessage} class, as shown in Figure
\ref{Fig:BetwMess}.  The information these two classes
need to contain is a \Robject{gUpdateDataMessage} object.  Thus, the
\Robject{gSendParentMessage} class has one slot, childUpdateDataMessage, and
this slot contains the \Robject{gUpdateDataMessage} object that was used to
update the child model.  Similarly, the \Robject{gSendChildMessage} class
has two slots, parentUpdateDataMessage and childName, where the
parentUpdateDataMessage slot contains the \Robject{gUpdateDataMessage} object
that was used to update the parent model and the childName slot contains the
name of the child model that is being updated (because a parent MVC can have
more than one child MVC).

Finally, the \Robject{gAskAncestorMessage} class, which inherits from the
\Robject{gMessage} class, is used to to ask an ancestor MVC for information
about its model.  This class is used when a child MVC is being created and it
needs information from more than just its immediate parent, i.e. it also needs
some information from its grandparent or a different ancestor.  The
\Robject{gAskAncestorMessage} class has three slots, from, type, and mData.
The from slot is the name of the MVC that asked for the information (i.e. where
this message came from), the type slot is the type of model to look for in an
ancestor, and the mData slot is the information being asked for in the model.

A generic function, \Rfunction{handleMessage}, is defined that is used to 
read message objects.  Methods for this generic function are defined in 
packages that use this package, such as \Rpackage{iSPlot} and
\Rpackage{iSNetwork}.

\section{Event-Callback Function Class}\label{Sec:EventFun}

Linking callback functions to events, such as a button click, a mouse movement
or a key press event, is what allows users to have an interactive
environment.  Then when an event happens something occurs in response
depending on the callback function linked to the event.  Now to create a
flexible environment we do not want a callback function that always does the
same thing when an event occurs.  For example, when a left button press event
occurs, we may sometimes want a point to be colored and at other times we may
want a point to be hidden.  To allow this flexibility, a class called
\Robject{gEventFun} was created.  By using this class, users are able to
change the response to an event.

The \Robject{gEventFun} class stores all of the information about a potential
callback function.  Having the callback function information stored in an 
object allows a user to connect events to callback functions.  The class 
definition for \Robject{gEventFun} is shown in Figure \ref{Fig:EventFun}.

\begin{figure}[ht]
  \begin{center}
    \scalebox{0.2}{\includegraphics{EventClass.jpg}}
    \caption{ The gEventFun Class. }
    \label{Fig:EventFun}
  \end{center}
\end{figure}

As shown in Fig \ref{Fig:EventFun}, the slots in the \Robject{gEventFun} class
are callFun, shortName, and preprocessFun.  The callFun slot is the callback
function, the shortName slot is a short description of the function, and the
preprocessFun slot is a list of all preprocessing functions that must be
called before the callback function.

The callback function that is stored in the callFun slot must take only one
parameter and the class of this parameter depends on the model class in
the active MVC object.  For example, for a data frame model, the parameter
is of class character, representing a row in the data frame.

The preprocessing functions that are stored in the preprocessFun slot
take no parameters.  The purpose of the preprocessing functions is to set some
variables that the callback function needs.  Thus, not all callback
functions need preprocessing functions and if no preprocessing functions
are needed, then this slot is set to NULL.  If the callback function sets the
color of a node in a graph, then a preprocessing function is needed to
determine what color is used.  So in this example, the preprocessing
function opens a color browser to allow the user to choose what the new
color is.

\section{Conclusions}\label{Sec:Conc}

The inheritance structure for the classes defined in this package are intended 
to be extensible.  If a user needs a model of a different type that is not 
currently represented, then a new model class can be defined that inherits 
from the \Robject{gModel} virtual class.  Similarly, new view and message 
classes can be created if new views or messages are needed by a user.

As mentioned in Section \ref{Sec:Overview} this package is intended to be a 
behind the scenes package that is used by other packages for its class and 
generic function definitions.  Currently, this package is used by the 
\Rpackage{iSPlot} and \Rpackage{iSNetwork} packages, and is extended by the
\Rpackage{BioMVCClass} package.

\end{document}